---
title: "gwang_hw2"
output: html_document
date: "2026-09-26"
---

Original code taken from:
- https://astrid-giugni.github.io/IDS_570_TAD/week-03-dictionaries.html
- https://astrid-giugni.github.io/IDS_570_TAD/week-04-representation.html

## Load packages

```{r packages, echo=FALSE, message=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(tidytext)
library(quanteda)
library(ggplot2)
library(tibble)
```

## Setup and preprocessing

Use exactly the same two texts as in Weeks 02–04:
• Circle of Commerce (Text A)
• Free Trade (Text B)

Use the same file paths and naming conventions introduced in the tutorials.

```{r setup}
txt_circle <- read_file("texts/A07594__Circle_of_Commerce.txt")
txt_free <- read_file("texts/B14801__Free_Trade.txt")
txt_third <- read_file("texts/A06785.txt")

texts <- tibble(
  doc_title = c("Text A", "Text B", "A06785"),
  text = c(txt_circle, txt_free, txt_third)
) %>%
  mutate(
    text_norm = text %>%
      str_replace_all("ſ", "s") %>% # normalize the “long s” (ſ) to s
      str_replace_all("\\s+", " ") %>%
      str_to_lower() # lowercase words only
  )

texts %>% select(doc_title, text_norm)
```

Stopwords were taken from the Week 04 tutorial along with tidytext's English stopword
list. Punctuation, numbers, and symbols were removed. Tokens are composed of words
only.

```{r stopwords}
data("stop_words")

custom_stopwords <- c(
  "vnto", "haue", "doo", "hath", "bee", "ye", "thee", "hee", "shall",
  "hast", "doe", "beene", "thereof", "thus"
)

all_stopwords <- bind_rows(
  stop_words %>% select(word),
  tibble(word = custom_stopwords)
) %>%
  distinct(word)

all_stopwords %>% slice_head(n = 10)
```

Create a tidy table with:
• one row per word token
• lowercase words
• stopwords removed
• Normalize the “long s” (ſ) to s, as we did in Week 03

```{r tokens}
clean_tokens <- texts %>%
  select(doc_title, text_norm) %>%
  unnest_tokens(word, text_norm, token = "words") %>%
  filter(str_detect(word, "^[a-z]+$")) %>% # alphabetic words only
  anti_join(all_stopwords, by = "word")

clean_tokens %>% slice_head(n = 10)
```

## I. Raw-Count Sentiment Analysis

Join the cleaned word tokens to the Bing sentiment lexicon. Use the Bing sentiment
lexicon to label words as positive or negative. Create a summary table with one row
per document that includes:

• number of positive words
• number of negative words
• net sentiment (positive − negative)

```{r raw-sentiment}
bing <- get_sentiments("bing") # create Bing table

# Bing labels some words ("envious", "enviously", and "enviousness") as both
# positive and negative. I've chosen to keep both labels such that each occurrence
# of these tokens contributes to both the positive and negative totals.
raw_sentiment_words <- clean_tokens %>%
  inner_join(bing, by = "word", relationship = "many-to-many") # keep tokens in dictionary

raw_sentiment_summary <- raw_sentiment_words %>%
  count(doc_title, sentiment, name = "word_count") %>% # count positive and negative words
  complete(
    doc_title = texts$doc_title,
    sentiment = c("positive", "negative"),
    fill = list(word_count = 0L)
  ) %>%
  pivot_wider(
    names_from = sentiment,
    values_from = word_count,
    names_prefix = "raw_",
    values_fill = 0 # replace NA values with 0
  ) %>% # one row per document
  mutate(raw_net_sentiment = raw_positive - raw_negative) %>% # calculate net sentiment
  filter(doc_title %in% c("Text A", "Text B")) %>%
  arrange(doc_title)

raw_sentiment_summary
```

## II. TF-IDF-Weighted Sentiment Analysis

### Setup and preprocessing

Use the same preprocessing decisions for both analyses. The R objects will be different
because Part I uses tidy data and Part II uses a quanteda DFM (to match the tutorials),
but the underlying text-cleaning decisions should be the same.

Then compute TF-IDF weights using dfm_tfidf(), as we did in Week 04.

```{r tfidf-dfm}
token_list <- clean_tokens %>%
  group_by(doc_title) %>%
  summarise(words = list(word), .groups = "drop") %>%
  tibble::deframe()

toks <- quanteda::as.tokens(token_list)
dfm_mat <- dfm(toks) # build dfm_mat, our raw-count matrix
tfidf_dfm <- dfm_tfidf(dfm_mat) # calculate tfidf_dfm to weigh words by frequency and distinctiveness

dfm_mat
```

### Convert TF-IDF weights from quanteda matrix into tidy format

The result of dfm_tfidf() is still a quanteda document-feature matrix. To join the words
to the Bing dictionary, we need to turn that matrix back into a regular data frame and 
then into a tidy format with one document-word combination per row.

Your resulting table should contain at least:
• the document name
• the word
• its TF-IDF weight

Remove rows where the TF-IDF weight is 0.

```{r tidy-tfidf}
tfidf_wide <- convert(tfidf_dfm, to = "data.frame") %>%
  rename(doc_title = doc_id) # resulting data table is wide (documents are rows and words are columns)

tfidf_tidy <- tfidf_wide %>% # turn into tidy table
  pivot_longer(
    cols = -doc_title,
    names_to = "word",
    values_to = "tf_idf"
  ) %>%
  filter(tf_idf > 0, doc_title %in% c("Text A", "Text B")) # remove rows with no weight
# Note: we are only analyzing Texts A & B here. The third was only used to calculate IDF.

tfidf_tidy %>% slice_head(n = 10)
```

### Inspect sentiment-bearing words with the highest TF-IDF weights

Join your tidy TF-IDF table to the Bing sentiment dictionary.

Once this is done, inspect the sentiment-bearing words with the highest TF-IDF weights in each
document. You will need these results to answer Question 3 in Part IV.

```{r top-tfidf-sentiment}
tfidf_sentiment_words <- tfidf_tidy %>%
  inner_join(bing, by = "word")

top_tfidf_sentiment_words <- tfidf_sentiment_words %>%
  group_by(doc_title) %>%
  slice_max(order_by = tf_idf, n = 20, with_ties = FALSE) %>% # filter, not recalculate, TF-IDF results
  ungroup() %>%
  arrange(doc_title, desc(tf_idf)) # arrange results by TF-IDF weight

print(top_tfidf_sentiment_words, n = 40) # words contributing the most weight to documents' positive and negative sentiment totals
```

### Sum positive and negative TF-IDF weights

Create a summary table with one row per document containing:
• total TF-IDF weight of positive words
• total TF-IDF weight of negative words
• net TF-IDF sentiment

Calculate:
net TF-IDF sentiment = positive TF-IDF weight − negative TF-IDF weight

```{r tfidf-summary}
tfidf_sentiment_summary <- tfidf_sentiment_words %>% # sum TF-IDF weights rather than counting rows/words
  group_by(doc_title, sentiment) %>%
  summarise(total_tf_idf = sum(tf_idf), .groups = "drop") %>% # one row per document & sentiment
  complete(
    doc_title = c("Text A", "Text B"),
    sentiment = c("positive", "negative"),
    fill = list(total_tf_idf = 0)
  ) %>%
  pivot_wider(
    names_from = sentiment,
    values_from = total_tf_idf,
    names_prefix = "tfidf_",
    values_fill = 0 # replace NAs with 0
  ) %>%
  mutate(tfidf_net_sentiment = tfidf_positive - tfidf_negative) %>%
  arrange(doc_title)

tfidf_sentiment_summary
```

## III. Compare Raw vs. TF-IDF Sentiment

Create a final table that includes:
• raw sentiment results
• TF-IDF–weighted sentiment results

Export this table as a .csv to be posted on Canvas.

```{r comparison}
sentiment_comparison <- raw_sentiment_summary %>%
  left_join(tfidf_sentiment_summary, by = "doc_title") %>% # combine summaries
  arrange(doc_title)

sentiment_comparison

write_csv(sentiment_comparison, "sentiment_comparison.csv")
```

## IV. Interpretive questions

Use the tables above to answer these questions in your Canvas response:

1. What changed between the raw-count and TF-IDF methods?
2. Why did TF-IDF alter the results?
3. Which specific words drove the changes? Use the top TF-IDF sentiment words
   table as evidence.
4. When would each method be more appropriate? What limitation does document
   length create for the raw-count sentiment measure?


## Optional challenge: sentiment concentration vs. diffusion

Using TF-IDF–weighted sentiment words:
1. Compute total TF-IDF weight of all sentiment-bearing words in each document (positive +
negative weights, without subtracting one from the other).
2. Identify the top 5 sentiment words per document.
3. Calculate the proportion of sentiment carried by those five words

```{r optional}
sentiment_concentration <- tfidf_sentiment_words %>%
  group_by(doc_title) %>%
  mutate(total_sentiment_tf_idf = sum(tf_idf)) %>%
  arrange(desc(tf_idf), .by_group = TRUE) %>%
  mutate(word_rank = row_number()) %>%
  summarise(
    total_sentiment_tf_idf = first(total_sentiment_tf_idf),
    top_five_words = str_c(word[word_rank <= 5], collapse = ", "),
    top_five_tf_idf = sum(tf_idf[word_rank <= 5]),
    top_five_proportion = top_five_tf_idf / total_sentiment_tf_idf,
    .groups = "drop"
  )

sentiment_concentration
```
